Assignment 2 report outline

Constraints (from assignment2_description.pdf):
- Single PDF per team (not one per member).
- Minimum 2 pages, maximum 6 pages.
- Must include the necessary data and plots to support each answer, not just prose.
- Submit by July 26, 2026, by email to apontec@mpip-mainz.mpg.de.
- State whether a grade (1-6) or a passed/failed confirmation is needed.

Header
- Matrikel-Nr 1, Matrikel-Nr 2.
- Grade vs. passed/failed choice.
(a few lines, not a full page)

Section 1. Setup summary (~0.25 page)
- One paragraph: three force fields (OPLS-AA/L+SPC, GROMOS96 43a1+SPC, GRAPPA on
  AMBER99SB-ILDN/TIP3P topology), each extended to 2 ns.
- Note what was kept identical across runs for a fair comparison: dt = 0.002 ps, Berendsen
  thermostat, constraint scheme, ion neutralization by total charge (see
  assignment2_method_clarifications.tex, Q3-Q5).

Section 2. General outcome -> Question 1 (~0.5 page)
- Number of integration steps for the 2 ns run (same for all three force fields).
- MAIN united-atom comparison = the PROTEIN-ATOM difference OPLS-AA vs. GROMOS
  (`outcome_df["protein_atoms_saved_vs_OPLS"]` in Part D.3). This is the clean number that
  isolates the united-atom effect (CH2/CH3 merged into single particles) and is the one to
  quote as the answer to "how many particles less".
- The full-SYSTEM atom count (`n_atoms_total`) is water-count-dependent: each force field is
  solvated and neutralized independently (and GRAPPA uses TIP3P vs. SPC), so the system total
  mixes the united-atom effect with run-to-run solvent differences. Report it only as context,
  with that caveat stated explicitly - do NOT quote the system-total difference as the
  united-atom answer.
- One-line sub-item for context: report the per-run water and ion counts (from the new
  system-composition table in Part D.3) so the reader can see where any system-total difference
  comes from.
- Table or bar chart: performance (ns/day) per force field and speed-up relative to OPLS-AA;
  connect the speed-up back to the protein-atom (and total-particle) reduction.

Section 3. Structural stability: RMSD and radius of gyration -> Question 2 (~1 page)
- RMSD and Rg overlay figure across the three force fields (whole-protein AND backbone/Ca
  selections side by side, `stability_summary` / the 2x2 overlay figure in the notebook).
- REQUIRED sensitivity comment (clarification Q1): explicitly state whether the whole-protein
  vs. backbone/Ca selection changes the stability conclusion (e.g. solvent-exposed side chains
  typically inflate whole-protein RMSD relative to backbone/Ca). One or two sentences; this is
  a graded clarification point, do not omit it.

Section 4. Flexibility: RMSF -> Question 3 (~1 page)
- RMSF overlay figure (per-residue, backbone) across the three force fields.
- Table of the residues with the largest RMSF change relative to OPLS-AA
  (`rmsf_diff_df`, top 5 per force field).
- One PyMOL snapshot per force field of interest, colored by b-factor (RMSF), to visually
  confirm affected regions are loops rather than secondary structure.
- Explanation of why GROMOS/GRAPPA might change flexibility at those residues, and how the
  hypothesis was cross-checked (PyMOL inspection, correlation with RMSD/SASA, caveat about
  single-trajectory statistics - clarification Q6/Q7).
- RMSF here is computed on the BACKBONE (consistent with the whole-protein vs. backbone/Ca
  selection discussion in Section 3); note this selection basis so the flexibility claims are
  comparable to the RMSD/Rg picture and not conflated with side-chain motion.

Section 5. Potential energy -> Question 4 (~0.5 page)
- Potential energy overlay figure across the three force fields (total Potential).
- Discussion: equilibration behavior, fluctuation magnitude; explicit caveat that absolute
  potential energy values are not comparable across force fields with different functional
  forms/reference points AND different system composition (water count/model) - focus on
  convergence and fluctuation, not absolute numbers.
- Support the "focus on relaxation, not absolute value" point with the additional, more
  defensible cross-force-field metric now computed in Part D.6 (protein-restricted /
  per-atom-normalized energy term), rather than comparing raw total Potential across FFs.
- V-rescale vs. Berendsen caveat (clarification Q4): note that the Berendsen thermostat was
  kept for strict consistency with the tutorial protocol, but that V-rescale would be the more
  rigorous choice for canonical (NVT) sampling in longer production runs. One sentence; this is
  an agreed clarification point and fits naturally in the energy/equilibration discussion (or
  in Section 1 setup - place it in whichever reads better, but include it).

Section 6. Solvent-accessible surface area -> Question 5 (~0.75 page)
- Total SASA + hydrophobic/hydrophilic split, per force field (the 1x3 small-multiples
  figure in the notebook).
- Hydrophobic-fraction table (`sasa_summary`) and discussion of any force field with a
  distinctly different exposed hydrophobic fraction.
- One PyMOL snapshot if a notable hydrophobic-patch difference is found (clarification Q8).

Section 7. Discussion: why compare force fields? -> Question 6 (~0.25 page)
- Short conceptual paragraph, not tied to a specific figure.

Section 8. Discussion: how can machine learning improve force fields? -> Question 7 (~0.25 page)
- Short conceptual paragraph, grounded in what was actually observed for GRAPPA vs. the two
  classical force fields in Sections 2-6.

References
- Original MD practical (Aponte-Santamaria / de Groot).
- Force field papers (OPLS-AA/L, GROMOS, GRAPPA - see References section of
  notebooks/assignment/practicalMD.ipynb).

Page budget check: Sections 2, 7, 8 are kept compact (table/bar chart or one short paragraph
each) so Sections 3-4 (the two questions needing the richest supporting figures) can use the
most space, while staying within the 6-page maximum.
